### 2017-11-17 Table_sum.tex
Added table comparisons

	\setlength\LTleft{-0.75in}
%	\setlength\LTright{-1in}
%\small
\footnotesize
    \begin{longtable}{C{2.1cm} C{2.1cm} C{2.3cm} C{1.5cm} C{2.7cm} C{2.2cm} }
    %\begin{longtable}{|C{2cm}|c|C{2.5cm}|C{3cm}|C{2.4cm}|c|}

		\hline
    	\textbf{Authors} & \textbf{Preprocessing} & \textbf{Segmentation} & \textbf{Features} & \textbf{Classification} & \textbf{Performance}  \\[1pt] \hline

    Ahirwar \emph{et al.}, 2012  &
    	- &
    	thresholding + granulometry, opening, morphological gradient, dilation, closing, thinning, spur removal &
    	- &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	- \vspace{0.6cm} \\
    	%\hline

    Anggraini \emph{et al.}, 2011 &
		- &
		thresholding + hole filling &	
		- &
		two (\emph{P.falciparum} infected and noninfected) + two life-cycle-stages &
		-
        \vspace{0.6cm}
		\\
    Arco \emph{et al.}, 2014 &
		- &
		adaptive thresholding + hole filling, closing, regional minima &	
		- &
		two (infected and noninfected) &
		-
		\vspace{0.6cm}
        \\
    Das \emph{et al.}, 2011 &
		- &
		marker controlled watershed &	
		opening, closing  &
		two (infected and noninfected) &
		-
		\vspace{0.6cm}
        \\

    Das \emph{et al.}, 2013 &
		- &
		marker controlled watershed &	
		- &
		three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
		-
		\vspace{0.6cm}
        \\

    Das \emph{et al.}, 2014 &
		- &
		marker controlled watershed &	
		- &
		three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
		-
		\vspace{0.6cm}
        \\

    Dave \emph{et al.}, 2017 &
		- &
		adaptive thresholding + erosion, dilation &	
		- &
		two (infected and noninfected) &
		-
		\vspace{0.6cm}
        \\

    Devi \emph{et al.}, 2017 &
		- &
		marker controlled watershed &	
		- &
		- &
		-
		\vspace{0.6cm}
        \\

        Dey \emph{et al.}, 2015 &
		- &
		opening &	
		- &
		- &
		-
		\vspace{0.6cm}
        \\

    Diaz \emph{et al.}, 2009 &
		- &
		inclusion tree &	
		- &
		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
		-
\vspace{0.6cm}
		\\

    Di Ruberto \emph{et al.}, 2002 &
		- &
		thresholding + granulometry, watershed  transform &	
		skeleton &
		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
		-
\vspace{0.6cm}
		\\

    Elter \emph{et al.}, 2011 &
		- &
		thresholding + black top-hat &	
		- &
		two (infected and noninfected) &
		-
		\vspace{0.6cm} \\

    Gonzalez-Betancourt \emph{et al.}, 2016  &
    	morphological filter, erosion-reconstruction, dilation-reconstruction, closing &
    	-  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Ghosh \emph{et al.}, 2011  &
    	- &
    	thresholding + opening, closing  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Kareem \emph{et al.}, 2012  &
    	dilation, erosion &
    	-  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Khan \emph{et al.}, 2011  &
    	- &
    	marker controlled watershed  &
    	- &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	- \vspace{0.6cm} \\

    Khan \emph{et al.}, 2014  &
    	- &
    	- &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Malihi \emph{et al.}, 2013  &
    	- &
    	-  &
    	area granulometry &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Mushabe \emph{et al.}, 2013  &
    	- &
    	thresholding + quali op.  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Oliveira \emph{et al.}, 2017  &
    	erosion &
    	-  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Reni \emph{et al.}, 2015  &
    	new morphological filtering &
    	-  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    Romero-Rondon \emph{et al.}, 2016  &
    	opening &
    	marker controlled watershed, erosion  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    Rosado \emph{et al.}, 2017  &
    	- &
    	adaptive thresholding + closing  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    Ross \emph{et al.}, 2006  &
    	area closing &
    	thresholding + granulometry, opening, reconstruction, morphological gradient, closing, thinning  &
    	erosion, dilation &
    	- &
    	- \vspace{0.6cm} \\

    Sheikhhosseini \emph{et al.}, 2013  &
    	hole filling, opening &
    	-  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    Savkare \emph{et al.}, 2011a  &
    	- &
    	watershed transform  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Savkare \emph{et al.}, 2011b  &
    	- &
    	thresholding + hole filling, watershed transform  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Savkare \emph{et al.}, 2015  &
    	- &
    	thresholding + watershed transform, erosion, dilation  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    {\v{S}}pringl, 2009  &
    	- &
    	thresholding + marker controlled watershed transform  &
    	area granulometry &
    	- &
    	- \vspace{0.6cm} \\

Somasekar \emph{et al.}, 2011  &
    	- &
    	- &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Somasekar \emph{et al.}, 2015  &
    	erosion, hole filling &
    	-  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Somasekar \emph{et al.}, 2017  &
    	- &
    	thresholding + erosion, closing  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Soni \emph{et al.}, 2011  &
    	- &
    	thresholding + granulometry, morphological gradient, dilation  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Sulistyawati \emph{et al.}, 2015  &
    	- &
    	blob analysis + erosion, dilation, opening, closing  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

 Tek \emph{et al.}, 2006  &
    	- &
    	top-hat, infinite reconstruction  &
    	area granulometry &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Tek \emph{et al.}, 2010  &
    	closing, granulometry &
    	thresholding + granulometry, area top-hat, closing  &
    	area granulometry &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) + 
    four life-cycle-stages for species &
    	- \vspace{0.6cm} \\

    Yunda \emph{et al.}, 2012  &
    	- &
    	thresholding + morphological gradient, erosion, dilation  &
    	- &
    	- &
    	- \\

		 \hline

	\caption{Summary of analysed methods: morphological operations used in the main phases of analysis, kind of classification and performance measures.} % needs to go inside longtable environment
	\label{tab:summary_table}
	\end{longtable}