### 2018-01-25 MDPI_template/Table.tex

%\setlength\LTleft{-0.75in}
%\setlength\LTright{-1in}
%\small
\footnotesize
\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
    \begin{longtable}{C{2.1cm} C{2.7cm} C{2.2cm} C{2.0cm} C{2.7cm} C{2.0cm} }
    %\begin{longtable}{|C{2cm}|c|C{2.5cm}|C{3cm}|C{2.4cm}|c|}

		\hline
    	\textbf{Authors} & \textbf{Preprocessing} & \textbf{Segmentation} & \textbf{MM/SE} & \textbf{Classification} & \textbf{Performance}  \\[1pt] \hline

%    Ahirwar \emph{et al.}, 2012  &
%    	- &
%    	thresholding + granulometry, opening, morphological gradient, dilation, closing, thinning, spur removal &
%    	gray, bin/disk of size depending on RBCs, diamond of size=1 &
%    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
%    	-  \\
%    	%\hline
%&   &  &  &  & \\
    Anggraini \emph{et al.}, 2011 &
		- &
		thresholding + hole filling &	
		gray, bin &
		two (\emph{P.falciparum} infected and noninfected) + two life-cycle-stages &
		SE=93\% SP=99\%
        %\vspace{0.6cm}
		\\
&   &  &  &  & \\
    Arco \emph{et al.}, 2014 &
		- &
		adaptive thresholding + hole filling, closing, regional minima &	
		gray, bin/disk &
		two (infected and noninfected) &
		Acc=96.46\%
		%\vspace{0.6cm}
        \\
        &   &  &  &  & \\
    Das \emph{et al.}, 2011 &
		- &
		marker controlled watershed + opening, closing &	
		gray, bin  &
		two (infected and noninfected) &
		Acc=88.77\%
		%\vspace{0.6cm}
        \\
        &   &  &  &  & \\
    Das \emph{et al.}, 2013 &
		- &
		marker controlled watershed &	
		gray &
	three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
		Acc=84\%
%		\vspace{0.6cm}
        \\
&   &  &  &  & \\
    Das \emph{et al.}, 2014 &
		- &
		marker controlled watershed &	
		gray &
	three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
		SE=99.72\% SP=84.39\%
		%\vspace{0.6cm}
        \\
&   &  &  &  & \\
    Dave \emph{et al.}, 2017 &
		- &
		adaptive thresholding + erosion, dilation &	
		bin &
		two (infected and noninfected) &
		Acc=97.83\% thin films, Acc=89.88\% thick films
		%\vspace{0.6cm}
        \\
&   &  &  &  & \\
    Devi \emph{et al.}, 2017 &
		- &
		marker controlled watershed &	
		gray &
		two (infected and noninfected)  &
		Acc=98.02\%
		%\vspace{0.6cm}
        \\
&   &  &  &  & \\
    Diaz \emph{et al.}, 2009 &
		- &
		inclusion tree &	
		gray &
		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
		SE=94\% SP=99.7\% for detection, SE=78.8\% SP=91.2\% for life-stages
%\vspace{0.6cm}
		\\
&   &  &  &  & \\
    Di Ruberto \emph{et al.}, 2002 &
		area closing, opening &
		thresholding + granulometry, watershed  transform + skeleton &	
		gray, bin/disk, flat and nonflat, with size depending on RBCs &
		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
		-
%\vspace{0.6cm}
		\\
&   &  &  &  & \\
    Elter \emph{et al.}, 2011 &
		- &
		thresholding + black top-hat, dilation &	
		gray/disk, nonflat with size=9 &
		two (infected and noninfected) &
		SE=97\%
%		\vspace{0.6cm}
\\
&   &  &  &  & \\
    Gonzalez-Betancourt \emph{et al.}, 2016  &
    	morphological filter, erosion-reconstruction, dilation-reconstruction, closing &
    	watershed  transform  &
    	gray/disk with size depending on RBCs &
    	- &
    	-
    %\vspace{0.6cm}
    \\
    &   &  &  &  & \\	
    Ghosh \emph{et al.}, 2011  &
    	- &
    	thresholding + opening, closing  &
    	bin &
    	two (\emph{P.vivax} infected and noninfected) &
    	-
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Kamreem \emph{et al.}, 2011, 2012  &
    	dilation, erosion &
    	-  &
    	gray/concentric ring, disk with size depending on RBCs  &
    	two (infected and noninfected) &
    	Acc=88\% SE=90\% SP=86\%
    %\vspace{0.6cm}
    \\	
&   &  &  &  & \\
    Khan \emph{et al.}, 2011  &
    	area closing &
    	thresholding + granulometry, opening, morphological reconstruction, gradient, dilation &
    	gray &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	Acc=81\% SE=85.5\%
    %\vspace{0.6cm}
    \\	
&   &  &  &  & \\
    Malihi \emph{et al.}, 2013  &
    	closing &
    	area granulometry  &
    	gray/disk with size depending on RBCs &
    	two (infected and noninfected) &
    	Acc=91\% SE=80\% SP=95.5\%
    %\vspace{0.6cm}
    \\	
&   &  &  &  & \\
    Mushabe \emph{et al.}, 2013  &
    	closing &
    	thresholding + granulometry, dilation, erosion  &
    	gray, bin/disk &
    	two (infected and noninfected) &
    	SE=98.5 SP=97.2\%
    %\vspace{0.6cm}
    \\	
&   &  &  &  & \\
    Oliveira \emph{et al.}, 2017  &
    	erosion &
    	gray, bin  &
    	- &
    	two (infected and noninfected) &
    	Acc=91\%
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Reni \emph{et al.}, 2015  &
    	new morphological filtering &
    	-  &
    	gray/anular ring, disk with size depending on RBCs &
    	- &
    	-
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Romero-Rondon \emph{et al.}, 2016  &
    	dilation, opening &
    	marker controlled watershed, erosion  &
    	gray, bin/disk with size depending on RBCs &
    	- &
    	-
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Rosado \emph{et al.}, 2017  &
    	- &
    	adaptive thresholding + closing  &
    	bin/elliptical with size=3 &
    	four (\emph{P.falciparum}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) +
    three life-cycle-stages for species &
    	SE=73.9-96.2\% SP=92.6-99.3\%
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Ross \emph{et al.}, 2006  &
    	area closing &
    	thresholding + granulometry, opening, reconstruction, morphological gradient, closing, thinning  &
        gray, bin/disk with size=6 and depending on RBCs, diamond with size=1 &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	SE=85\% for detection, Acc=73\% for classification
    	 %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Savkare \emph{et al.}, 2011a  &
    	- &
    	thresholding + hole filling, watershed transform  &
    	bin &
    	two (infected and noninfected) &
    	-
%    \vspace{0.6cm}
    \\
&   &  &  &  & \\
    Savkare \emph{et al.}, 2011b  &
    	- &
    	thresholding + hole filling, watershed transform  &
    	bin &
    	two (infected and noninfected) &
    	SE=93.12\% SP=93.17\%
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Savkare \emph{et al.}, 2015  &
    	- &
    	thresholding + watershed transform, erosion, dilation  &
    	bin/disk with size=2 &
    	two (infected and noninfected) &
    	Acc=95.5\%
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Sheikhhosseini \emph{et al.}, 2013  &
    	hole filling &
    	thresholding + hole filling, opening  &
    	bin &
    	two (infected and noninfected) &
    	Acc=97.25\% SE=82.21\% SP=98.02\%
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Somasekar \emph{et al.}, 2015  &
    	erosion &
    	fuzzy C-means clustering + erosion, hole filling  &
    	bin/square with size=3 &
    	two (infected and noninfected) &
    	SE=98\% SP=93.3\%
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Somasekar \emph{et al.}, 2017  &
    	- &
    	thresholding + erosion, closing, hole filling  &
    	bin/square with size=3 &
    	two (infected and noninfected) &
    	average DSC=0.8
    %\vspace{0.6cm}
    \\
%&   &  &  &  & \\
%    Soni \emph{et al.}, 2011  &
%    	- &
%    	thresholding + granulometry, morphological gradient, dilation  &
%    	gray, bin/disk, flat and nonflat, vertical, horizontal &
%    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
%    	SE=98\% for detection
%    %\vspace{0.6cm}
%    \\
&   &  &  &  & \\
{\v{S}}pringl, 2009  &
    	closing &
    	thresholding + marker controlled watershed transform, hole filling, dilation, opening, erosion  &
    	gray, bin/disk with size depending on RBCs &
    	two (infected and noninfected) &
    	AUC=0.98
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Sulistyawati \emph{et al.}, 2015  &
    	- &
    	blob analysis + erosion, dilation, opening, closing, hole filling  &
    	bin &
    	two (infected and noninfected) &
    	Acc=99.39\%
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
 Tek \emph{et al.}, 2006  &
    	- &
    	top-hat, infinite reconstruction, area granulometry  &
    	gray, bin &
    	two (infected and noninfected) &
    	SE=74\% SP=98\%
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Tek \emph{et al.}, 2010  &
    	closing, granulometry &
    	thresholding + granulometry, area top-hat, closing, area granulometry &
    	gray/disk with size depending on RBCs &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) +
    four life-cycle-stages for species &
    	SE=72\% SP=98\%
    %\vspace{0.6cm}
    \\
&   &  &  &  & \\
    Yunda \emph{et al.}, 2012  &
    	- &
    	thresholding + morphological gradient, erosion, dilation  &
    	bin &
    	three (\emph{P.falciparum}, \emph{P.vivax} infected, and noninfected) + two life-cycle-stages for \emph{P.falciparum}&
    	SE=77.19\% \\

		 \hline

	\caption{Summary of analysed methods: morphological operations used in the main phases of analysis, type of MM (gray or binary)/type and size of SE (if reported), kind of classification and performance measures (Sensitivity, Specificity, Accuracy, if reported).} % needs to go inside longtable environment
	\label{tab:summary_table}
	\end{longtable}